---
title: "iteration_and_listcols"
author: "Xiaoni Xu"
date: "2024-10-29"
output: html_document
---


```{r, message = FALSE}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

## **Lists**

```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Justin")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
# not the best way
```

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))
l

l$mat # Jeff hates it

l[["mat"]][1, 3] # Jeff likes it (insert hotline bling meme)
```

Make a list that's hopefully a bit more useful.
```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norm[["b"]]
```

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}

# click shift+alt at the same time and select multiple rows, to type multiple rows at once
```

```{r}
mean_and_sd(list_norm[[1]])
mean_and_sd(list_norm[[2]])
mean_and_sd(list_norm[[3]])
mean_and_sd(list_norm[[4]])
```

## **Use a for loop**
```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norm[[i]])
}

output
```

## Do the same thing but with `map` instead
```{r}
output = map(list_norm, mean_and_sd) %>% 
  bind_rows()

output = map_dfr(list_norm, mean_and_sd)
```

do a couple other things
```{r}
output = map_dbl(list_norm, IQR)
```

## List columns

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm # much match the number of things in the `name` column
  )

listcol_df

listcol_df %>% 
  filter(name %in% c("a", "b"))

listcol_df %>% 
  select(-samp)
```
```{r}
listcol_df[["samp"]][["a"]]
```

Compute mean and sd

```{r}
mean_and_sd(listcol_df[["samp"]][["a"]])
mean_and_sd(listcol_df[["samp"]][["b"]])

map(listcol_df[["samp"]], mean_and_sd)
```
add a list column
```{r}
listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR))

listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)) %>% 
  select(-samp) %>% 
  unnest(output)
```

## **Revisiting NSDUH**

```{r}

nsduh_table_format <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}
```

We can use this function to import three tables using the next code chunk, which downloads and extracts the page HTML and then iterates over table numbers. The results are combined using `bind_rows()`. Note that, because this version of our function doesn’t include `table_name`, that information is lost for now.
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

output = vector("list", 3)

for (i in c(1, 4, 5)) {
  output[[i]] = nsduh_html(nsduh_html, i)
}

nsduh_results = bind_rows(output)
```


```{r}
nsduh_table_format(html = nsduh_html, table_num = 1)
nsduh_table_format(html = nsduh_html, table_num = 4)
nsduh_table_format(html = nsduh_html, table_num = 5)
```


```{r}
# create a different column for every drug, and then create a table for something???
nsduh_df = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_n = c(1, 4, 5)
  ) %>% 
  mutate(table = map(table_n, nsduh_table_format, html = nsduh_html)) %>% 
  unnest(table)

# just a plain old dataframe, and now we can do tidy things

nsduh_df %>% 
  filter(State == "New York")


```

## **Operations on nested data**

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

Create a list column
```{r}
weather_nest = 
  nest(weather_df, data = date:tmin)
```

```{r}
weather_nest[["data"]][[1]]
```

Suppose we want to fit the simple linear regression relating tmax to tmin for each station-specific data frame. First I’ll write a quick function that takes a data frame as the sole argument to fit this model.

```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
}
```

Let’s make sure this works on a single data frame.
```{r}
weather_lm(weather_nest$data[[1]])
```

```{r}
weather_nest %>% 
  mutate(model_fit = map(data, \(x) lm(tmax ~ tmin, data = x)))
```

